Borel regularity for ODEs and Thimble Integrals

### 2024-01-17 exp-resurgence.tex
Sync before meeting

Once you have the formal solution $\series{\Phi}$, you might try to get an actual solution by applying {\em Borel summation}, which turns a formal power series into a function asymptotic to it. Borel summation works in three steps.
\begin{enumerate}
\item Thinking of $z$ as a ``frequency variable,'' we take the Borel transform (also known as formal inverse Laplace transform) of $\series{\Phi}$, producing a formal power series $\series{\phi}$ in a new ``position variable'' $\zeta$.
\item If $\series{\phi}$ has a positive radius of convergence we say $\series{\Phi}$ is {\em $1$-Gevrey}. We sum $\series{\phi}$ to get a holomorphic function $\hat{\phi}$ on a neighborhood of $\zeta = 0$. Then, by analytic continuation, we expand the domain of $\hat{\phi}$ to a Riemann surface $B$ with a distinguished 1-form $\lambda$---the continuation of $d\zeta$. %\textcolor{orange}{[Nikita has a complementary picture where the Borel plane is the cotangent fiber? Ask more about this.]}
\item Furthermore, if $\hat{\phi}$ grows slowly enough along an infinite ray $b + e^{i\theta}[0, \infty)$ \textcolor{magenta}{[change, explain, or link to notation]} its Laplace transform $\laplace_b^\theta \hat{\phi}$ (see Section \ref{laplace:ordinary}) turns out to be a holomorphic function of $z$, well-defined on some sector of the frequency plane. In this case, we say $\tilde{\Phi}$ is {\em Borel-summable}, and we call $\hat{\Phi}:=\laplace_b^\theta \hat{\phi}$ its {\em Borel sum} at $b$. 
\end{enumerate}
The series $\series{\Phi}$ and its Borel sum $\hat{\Phi}$ have a special relationship, which is best described in the language of {\em Gevrey asymptoticity}. %\textcolor{magenta}{[Do we need to extend these definitions, and the associated theorems, to cover slight series? Hopefully it wouldn't be too hard, but it would be a pain.]}
\begin{defn}
On an open, possibly bounded sector $\Omega$ around $\infty$, a power series
\begin{equation}\label{series:asymp-defn}
\frac{a_1}{z} + \frac{a_2}{z^2} + \frac{a_3}{z^3} + \frac{a_4}{z^4} + \ldots
\end{equation}
is {\em asymptotic} to a holomorphic function $F$ if
\[ \left|F - \left(\frac{a_1}{z} + \frac{a_2}{z^2} + \ldots + \frac{a_{n-1}}{z^{n-1}} \right) \right| \in o_{z \to \infty}\big(|z|^{-n}\big) \]
over all orders $n$. We'll use the symbol $\sim$ to denote the asymptotic of $F$, namely we'll write \[F\sim \sum_{n\geq 1} a_n z^{-n}.\]
\end{defn}
If two power series of the form \eqref{series:asymp-defn} are asymptotic to the same function, they must be equal; this can be deduced from the triangle inequality. The power series which is asymptotic to a given function on a given open sector, if it exists, is therefore an intrinsic property of the function: it's the function's {\em asymptotic expansion} on that sector. The functions that have asymptotic expansions on a given sector form a ring~\cite[Section~A.4]{nikolaev2023existence}, and the map that sends such a function to its asymptotic expansion is a ring homomorphism.

@@

The central goal of this paper is to explain, from this perspective, why Borel summation works well for the two kinds of problems exemplified in Section~\ref{intro:summation}.
\begin{enumerate}
\item Solving a linear ODE of single level~$1$ \cite[Section 2.1]{EcalleIII}\textcolor{DarkTurquoise}{\cite[Section~5.2.2.1]{diverg-resurg-iii}}. One example is Equation~\eqref{eqn:bessel-rescaled-1}. The general example is the equation $\mathcal{P} \Phi = 0$ given by a differential operator of the form
\[ \mathcal{P} = P(\partial_z) + \frac{1}{z} Q(\partial_z) + \frac{1}{z^2}\big[ R_0(z^{-1}) + R_1(z^{-1})\,\partial_z + \ldots + R_{d-1}(z^{-1})\,\partial_z^{d-1} \big], \]
where $P$ is a monic degree-$d$ polynomial, $Q$ is a degree-$(d-1)$ polynomial, and $R_0, \ldots, R_{d-1}$ are holomorphic functions in some disk $|z|>A$ around infinity. We'll restrict our attention to the case where the roots of $P$ are simple (see Section \ref{borel_reg-ODE}).

@@

\begin{remark}
Our approach of solving regular singular Volterra equation is analogous to the one that Braaksma used to solve a different class of non-linear ODEs and difference equations whose coefficients are written as Laplace transforms~\cite{braaksma2006laplace}. \textcolor{orange}{[Let's do more compare-and-contrast! Our problem space is sort of orthogonal to Braaksma's, but we use the same solution space, so it seems like there could be a cool interaction there.]}
\end{remark}

@@

In Section~\ref{borel_reg-ODE}, we show that we're all finding the same frame. The uniqueness part of \cite[Theorem~4]{reg-sing-volterra} shows that \'{E}calle's formal solutions sum to our analytic ones, and the properties of the Laplace transform guarantee that our solutions, cast into in the frequency domain, are asymptotic to Poincar\'{e}'s.
\end{remark} 
%
%
%
\subsection{Borel regularity for thimble integrals}
%
Thimble integrals have been studied from different perspectives: in physics, they play an important technical role in quantum mechanics, where infinite-dimensional exponential integrals are supposed to give the expectation values of observable quantities \cite{dunne-unsal2,dunne-unsal,Fauvet_Menous_Queva,Tanizaki:2014tua}. In this context, physicists often use Borel summation, resurgence and related techniques to assign values to these integrals going beyond their pertubative expansion. For instance, in complex Chern--Simons theory, one could decompose the path integral in thimble integrals and arguing as in the finite dimesional set-up to study the Witten--Reshetikhin--Turaev invariants of $3$ manifolds \cite{gukov-marino-purtrov-resurgence,Witten}. \textcolor{orange}{Take a look at Witten's paper on knots} More generally, we could of a problem that is known to admit a thimble integral solution but hard to be defined explicitly. However, we know the perturbative expansion of this expected solution; then Borel regularity tells us that the Borel sum of the pertubative series is the solution we're looking for. In fact, the Laplace transform ray (namely the ray along which the Laplace trasnform is defined) is often a ``collapsed'' thimble.\textcolor{Maroon}{This is the bit about turning a perturbative series back into a thimble integral, taking advantage of the fact that the Laplace transform ray is often a ``collapsed'' thimble.}

In the algebraic geometric set-up, namely when $X$ is an $N$-dimensional alegbraic variety over $\C$ and $f$ is a proper map $f\colon X\to\C$, thimble integrals are known as period integrals \cite{deligne2007singularites,Maxim_lectures,pham}.\footnote{In particular, they find application in mirror symmetry for Fano varieties as they encode the Gromov--Witten invariants.\textcolor{orange}{add references---Maxim's mirror symmetry talk, Varchenko}.} In fact, the thimbles represent classes in the homology $H^{B,\theta}_N(X,f)$ which are relative to the preimage of $e^{i\theta}\infty$ under the funtion $f$, for a generic direction $\theta$.\footnote{The relative homology $H_i^{\theta}(X,f)$ is defined as the limit as $c\to e^{i\theta}\infty$, of $H_i^{\theta}(X,f^{-1}(S_c^+))$, where $S_c^+=\lbrace \zeta\in\C \vert \Re \zeta> c\rbrace$. Generic directions correspond to $\theta\neq \arg (\alpha-\beta)$. } Furthermore, if $f$ has non-degenerate critical points, they form a basis for $H_N^{B,\theta}(X,f)$. Then varying $\theta$, $H_N^{B,\theta}(X,f)$ forms a local system over $S^1=\R/2\pi \Z$ singular at $\theta=\arg(\alpha-\beta)$, and whose monodromy can be computed by Picard--Lefschetz formula (we refer to \cite{pham,Arnold}).

\textcolor{orange}{thimble integral as solution of ODE with irregular singularity and the Stokes phenomena. }

Indeed, the Stokes data of an irregular singular connection can be computed by the Picard--Lefschetz formula (see~\cite[Section 3.3, Part II]{pham}).\textcolor{orange}{add references}  This can be also considered a consequence of our Borel regularity result when $f$ is a polynomial. For instance, special functions are classical examples of thimble integrals (with polynomial $f$), thus the Borel summability properties of their asymptotics can be studied either analytically (through the \textit{thimble projection reasoning} see the discussion in Section~\ref{sec:examples}) or geometrically (in terms of a certain Riemann--Hilbert problem \cite{Maxim_slide_ERC}\cite[Section 6.2]{kontsevich2022analyticity}).

In the $N$-dimensional analog of our set-up \ref{borel-reg:explanatory-power}\footnote{In fact, $f$ might satisfy milder assumptions (see for instance~\cite[Section 1.2.2]{mistegard_phdthesis}).}, the thimbles are actually steepest descent paths from the critical points; thus the asymptotic behaviour as $z\to e^{i\theta}\infty$ can be studied using the steepest descent method \cite{andersen2020resurgence,delabaere-howls,delabaere_dillinger_pham,Delabaere-Pham99,dingle1973asymptotic,Malgrange22,Pham83}. However, the asymptotics of thimble integrals is typically a formal power series, and choosing a suitable resummation technique one should be able to recover the original function. This raises the question we address in this paper, namely to understand why Borel summation is effective for $1$-dimensional thimbles integrals. In a nutshell, our result follows from the fact that thimble integrals are generalized Laplace transforms, hence they are Borel regular functions. In addition. our approach wants to focus on the role of the position domain and it emphasizes how the geometry of the Laplace transform is the natural framework to describe Borel regularity for thimbles integrals (see Fenyes's lecture \cite{Fenyes-ihes-lecture}).

@@

\textcolor{orange}{hyperasymptotics \cite{Berry_Howls,Berry1991,Howls97,Howls}}
\color{DarkTurquoise}
These questions have been classically addressed by Pham \cite[Section 3.3]{pham}, Malgrange \cite[Theorem 6]{Malgrange22}, Berry and Howls \cite{Berry_Howls} and Sternin and Shatalov \cite[Theorem 3.9 and 3.10]{sternin1995borel}. In addition, the study of thimble integrals plays an important technical role in quantum mechanics, where infinite-dimensional exponential integrals are supposed to give the expectation values of observable quantities. Physicists often use Borel summation, resurgence and related techniques to assign values to these integrals \cite{costin_kruskal,pham1988resurgence}.
\color{black}
%Our approach wants to focus on the role of the position domain and it emphasizes how the geometry of the Laplace transform is the natural framework to describe Borel regularity for thimbles integrals (see Fenyes's lecture \cite{Fenyes-ihes-lecture}). In a nutshell, we use that thimble integrals are generalized Laplace transforms, to show they are indeed Borel regular functions.  
%     
\section{The Laplace and Borel transforms}\label{sec:Laplace-Borel-general}
\subsection{The geometry of the Laplace transform}
Classically, the Laplace transform turns functions on the position domain into functions on the frequency domain. In the study of Borel summation and resurgence, it's useful to see the position domain as a {\em translation surface} $B$, and the frequency domain as one of its cotangent spaces. Roughly speaking, the Laplace transform lifts holomorphic functions on $B$ to holomorphic functions on $T^* B$.
%
\subsubsection{Translation surfaces, briefly}
%
A translation surface is a Riemann surface $B$ carrying a holomorphic $1$-form $\lambda$~\cite{zorich2006flat}. A translation chart is a local coordinate $\zeta$ with $d\zeta = \lambda$. The standard metric on $\C$ pulls back along translation charts to a flat metric on $B$, with a conical singularity of angle $2\pi n$ wherever $\lambda$ has a zero of order $n-1 > 0$. We'll require $B$ to be finite-type and $\lambda$ to have a pole at each puncture. This kind of translation surface has a ``cylindrical end'' (see Figure \ref{fig:translation_surface}) at each puncture where $\lambda$ has order $-1$, and a ``$|2n|$-planar end'' (see Figure \ref{fig:translation_surface}) at each puncture where $\lambda$ has order $n-1 < -1$~\cite[Section 2.5]{gupta2013meromorphic} (or cite Aaron's article, which will hopefully present the same background in the translation surface context).
\begin{figure}[ht]
    \centering
    %\includegraphics{}
    \caption{On the left an example of translation surface with ``cylindrical end''. On the right, an example of translation surface with  ``$|2n|$-planar end''. }
    \label{fig:translation_surface}
\end{figure}
%
\subsubsection{Direction}\label{transl:dir}
%
The translation structure gives $B$ a notion of direction as well as distance. Away from the zeros of $\lambda$, which we'll call {\em branch points}, we can talk about moving upward, rightward, or at any angle, just as we would on $\C$. At a branch point of cone angle $2\pi n$, we can also talk about moving upward, rightward, or at any angle in $\R/2\pi\Z$, but here there are $n$ directions that fit each description. To make this more concrete, note that around any point $b \in B$, there's a unique holomorphic function $\zeta_b$ that vanishes at $b$ and has $d\zeta_b = \lambda$. \textcolor{VioletRed}{[If we define ``translation parameter'' earlier, we can say:] there's a unique translation parameter $\zeta_b$ that vanishes at $b$.} This function is a translation chart when $b$ is an ordinary point, and an $n$-fold branched covering when $b$ is a branch point of cone angle $2\pi n$. In either case, $\zeta_b \in e^{i\theta} [0, \infty)$ is a ray or a set of rays leaving $b$ at angle $\theta \in \R/2\pi\Z$.
%

@@

\item With respect to the flat metric, \textcolor{red}{$f$ is uniformly of exponential type $\Lambda$ along $\Gamma_b^\theta$, and is locally integrable throughout.}\footnote{Recall that a function is of exponential type $\Lambda$ if for every $\varepsilon>0$ there is a constant $A_\varepsilon$ (which depends on $\varepsilon$) such that $|f(x)|\leq A_\varepsilon e^{\Lambda+\varepsilon} $. We instead require that there exists a uniform constant $A$ such that for every $\varepsilon>0$ the bound holds.} Following \cite{reg-sing-volterra}, we require $f\in\singexp{\sigma}{\Lambda}(\Omega_b)$ with $\sigma>-1$.\footnote{The condition $\sigma>-1$ guarantees the integrability at $\zeta=b$.} The set $\Omega_b$ is a tubular neighbourhood of $\Gamma_b^\theta$. 

@@

\color{RoyalBlue}{$\singexpalg{\sigma}(\Omega_\alpha^\Lambda)$}
\color{Maroon}
Suppose $\Omega_\alpha$ is an open sector with $\zeta = \alpha$ at its tip, and an opening angle of $\pi$ or less. In this case, for any $\Lambda \in \R$, let \textcolor{magenta}{$\widehat{\Omega}_\alpha^\Lambda$} $\dualsector{\alpha}{\lambda}$ be the union of the half-planes $\Re(e^{i\theta} z) > \Lambda$ over all angles $\theta$ in the opening of $\Omega_\alpha$. Then, let $\dualsingexp{\sigma}{\Lambda}(\Omega_\alpha)$ be the space of holomorphic functions $F$ on $\widehat{\Omega}_\alpha^\Lambda$ with $|F| \lesssim \Delta^\sigma$, where $\Delta$ is the distance to the boundary of $\widehat{\Omega}_\alpha^\Lambda$. The norm $\|F\|_{\sigma, \Lambda} = \sup_{\widehat{\Omega}_\alpha^\Lambda} \Delta^{-\sigma} |F|$ turns $\dualsingexp{\sigma}{\Lambda}(\Omega_\alpha)$ into a Banach space. \textcolor{orange}{Cite \cite{sternin1995borel} better for the theorem.}
\begin{prop}[following~\cite{sternin1995borel}]\label{prop:laplace-cont}
Let $\Omega_\alpha$ be an open sector with $\zeta = \alpha$ at its tip, and an opening angle of $\pi$ or less. For any $\sigma > 0$ and $\Lambda \ge 0$, and any angle $\theta$ in the opening of $\Omega_\alpha$, the Laplace transform $\laplace_{\zeta_\alpha, 0}^\theta$ is a continuous map $\singexp{\sigma-1}{\Lambda}(\Omega_\alpha) \to \dualsingexp{-\sigma}{\Lambda}(\Omega_\alpha)$, with a norm of at most $\Gamma(\sigma)$.
\end{prop}
\begin{proof}
Given some $f \in \singexp{\sigma-1}{\Lambda}(\Omega_\alpha)$, we compute
\begin{align*}
|\laplace_{\zeta_\alpha, 0}^\theta f| & = \left| \int_{\Gamma_{\zeta, \alpha}^\theta} e^{-z\zeta_\alpha} f\,d\zeta_\alpha \right| \\
& \le \int_{\Gamma_{\zeta, \alpha}^\theta} e^{-\Re(z\zeta_\alpha)} |\zeta_\alpha|^{\sigma-1} e^{\Lambda |\zeta_\alpha|} \|f\|_{\sigma-1, \Lambda}\,|d\zeta_\alpha| \\
& \le \int_{\Gamma_{\zeta, \alpha}^\theta} e^{(\Lambda - c_{z, \theta}|z|)|\zeta_\alpha|} |\zeta_\alpha|^{\sigma-1} \|f\|_{\sigma-1, \Lambda}\,|d\zeta_\alpha|,
\end{align*}
where $c_{z, \theta}$ is the cosine of $\arg(z) + \theta$. When $|z|$ is large and $c_{z, \theta}$ is positive, the integrand shrinks exponentially as $|\zeta|$ grows. This shows that for each angle $\theta$ in the opening of $\Omega_\alpha$, the integral defining $\laplace_{\zeta_\alpha, 0}^\theta f$ converges in some region of $\widehat{\Omega}_\alpha^\Lambda$. It also shows that for different angles $\theta$, the functions $\laplace_{\zeta_\alpha, 0}^\theta f$ match where their domains overlap. We can therefore glue these functions together into one big Laplace transform of $f$, defined at large values of $|z|$ across the whole opening angle of $\widehat{\Omega}_\alpha^\Lambda$.

>>>>

Once you have the formal solution $\series{\Phi}$, you might try to get an actual solution by applying {\em Borel summation}, which turns a formal power series into a function asymptotic to it. Borel summation works in three steps.
\begin{enumerate}
\item Thinking of $z$ as a ``frequency variable,'' we take the Borel transform (also known as formal inverse Laplace transform) of $\series{\Phi}$, producing a formal power series $\series{\phi}$ in a new ``position variable'' $\zeta$.
\item If $\series{\phi}$ has a positive radius of convergence we say $\series{\Phi}$ is {\em $1$-Gevrey}. We sum $\series{\phi}$ to get a holomorphic function $\hat{\phi}$ on a neighborhood of $\zeta = 0$. Then, by analytic continuation, we expand the domain of $\hat{\phi}$ to a Riemann surface $B$ with a distinguished 1-form $\lambda$---the continuation of $d\zeta$. %\textcolor{orange}{[Nikita has a complementary picture where the Borel plane is the cotangent fiber? Ask more about this.]}
\item Furthermore, if $\hat{\phi}$ grows slowly enough along an infinite ray $b + e^{i\theta}[0, \infty)$ \textcolor{magenta}{[change, explain, or link to notation]} its Laplace transform $\laplace_b^\theta \hat{\phi}$ (see Section \ref{laplace:ordinary}) turns out to be a holomorphic function of $z$, well-defined on some sector of the frequency plane. In this case, we say $\tilde{\Phi}$ is {\em Borel-summable}, and we call $\hat{\Phi}:=\laplace_b^\theta \hat{\phi}$ its {\em Borel sum} at $b$. 
\end{enumerate}
The series $\series{\Phi}$ and its Borel sum $\hat{\Phi}$ have a special relationship, which is best described in the language of {\em Gevrey asymptoticity}. %\textcolor{magenta}{[Do we need to extend these definitions, and the associated theorems, to cover slight series? Hopefully it wouldn't be too hard, but it would be a pain.]}
\begin{defn}
On an open, possibly bounded sector $\Omega$ around $\infty$, a power series
\begin{equation}\label{series:asymp-defn}
\frac{a_1}{z} + \frac{a_2}{z^2} + \frac{a_3}{z^3} + \frac{a_4}{z^4} + \ldots
\end{equation}
is {\em asymptotic} to a holomorphic function $F$ if
\[ \left|F - \left(\frac{a_1}{z} + \frac{a_2}{z^2} + \ldots + \frac{a_{n-1}}{z^{n-1}} \right) \right| \in o_{z \to \infty}\big(|z|^{-n}\big) \]
over all orders $n$. We'll use the symbol $\sim$ to denote the asymptotic of $F$, namely we'll write \[F\sim \sum_{n\geq 1} a_n z^{-n}.\]
\end{defn}
If two power series of the form \eqref{series:asymp-defn} are asymptotic to the same function, they must be equal; this can be deduced from the triangle inequality. The power series which is asymptotic to a given function on a given open sector, if it exists, is therefore an intrinsic property of the function: it's the function's {\em asymptotic expansion} on that sector. The functions that have asymptotic expansions on a given sector form a ring~\cite[Section~A.4]{nikolaev2023existence}, and the map that sends such a function to its asymptotic expansion is a ring homomorphism. We'll denote such map with the symbol $\aexp$. More generally, when the asymptotic expansion is taken in the direction $\theta$ at infinity, i.e. as $z\to e^{i\theta}\infty$, we'll write $\aexp^\theta$.

@@

The central goal of this paper is to explain, from this perspective, why Borel summation works well for the two kinds of problems exemplified in Section~\ref{intro:summation}.
\begin{enumerate}
\item Solving a linear ODE of single level~$1$ \cite[Section 2.1]{EcalleIII}\cite[Section~5.2.2.1]{diverg-resurg-iii}. One example is Equation~\eqref{eqn:bessel-rescaled-1}. The general example is the equation $\mathcal{P} \Phi = 0$ given by a differential operator of the form
\[ \mathcal{P} = P(\partial_z) + \frac{1}{z} Q(\partial_z) + \frac{1}{z^2}\big[ R_0(z^{-1}) + R_1(z^{-1})\,\partial_z + \ldots + R_{d-1}(z^{-1})\,\partial_z^{d-1} \big], \]
where $P$ is a monic degree-$d$ polynomial, $Q$ is a degree-$(d-1)$ polynomial, and $R_0, \ldots, R_{d-1}$ are holomorphic functions in some disk $|z|>A$ around infinity. We'll restrict our attention to the case where the roots of $P$ are simple (see Section \ref{borel_reg-ODE}).

@@

\begin{remark}
Our approach of solving regular singular Volterra equation---based on the contraction mapping theorem for suitable Banach spaces \cite{reg-sing-volterra}---is analogous to the one that Braaksma used to solve a different class of non-linear ODEs and difference equations whose coefficients are written as Laplace transforms~\cite{braaksma2006laplace}. What distinguishs our result from the one of Braaksma is to consider the Laplace transform of functions with integrable fractional power singuarities, and even if our result has been proved for linear level~1 ODEs, we allow holomorphic coefficents that are not necesserly Laplace transforms.

We leave a possible generalization of our result to other classes of ODEs and difference equations for further pubblications. \textcolor{orange}{[Let's do more compare-and-contrast! Our problem space is sort of orthogonal to Braaksma's, but we use the same solution space, so it seems like there could be a cool interaction there.]}
\end{remark}

@@

In Section~\ref{borel_reg-ODE}, we show that we're all finding the same frame. The uniqueness part of \cite[Theorem~4]{reg-sing-volterra} shows that \'{E}calle's formal solutions sum to our analytic ones, and the properties of the Laplace transform guarantee that our solutions, cast into in the frequency domain, are asymptotic to Poincar\'{e}'s.
\end{remark} 
%
%
%
\subsection{Borel regularity for thimble integrals}
%
Thimble integrals have been studied from different perspectives: in physics, they play an important technical role in quantum mechanics, where infinite-dimensional exponential integrals are supposed to give the expectation values of observable quantities \cite{dunne-unsal2,dunne-unsal,Fauvet_Menous_Queva,Tanizaki:2014tua}. In this context, physicists often use Borel summation, resurgence and related techniques to assign values to these integrals going beyond their pertubative expansion (see also \cite{Berry_Howls,Berry1991,costin_kruskal,Howls97,Howls,pham1988resurgence}). For instance, in complex Chern--Simons theory, one could decompose the path integral in thimble integrals and arguing as in the finite dimesional set-up to study the Witten--Reshetikhin--Turaev invariants of $3$ manifolds \cite{gukov-marino-purtrov-resurgence,Witten}. \textcolor{orange}{Take a look at Witten's paper on knots} More generally, we could think of a problem that is known to admit a thimble integral solution but that's hard to be defined explicitly. However, we know the perturbative expansion of the expected solution; then Borel regularity tells us that the Borel sum of the pertubative series is the solution we're looking for. In fact, the Laplace transform ray (namely the ray along which the Laplace trasnform is defined) is often a ``collapsed'' thimble.\textcolor{Maroon}{This is the bit about turning a perturbative series back into a thimble integral, taking advantage of the fact that the Laplace transform ray is often a ``collapsed'' thimble.}

In the algebraic geometric set-up, namely when $X$ is an $N$-dimensional alegbraic variety over $\C$ and $f$ is a proper map $f\colon X\to\C$, thimble integrals are known as period integrals \cite{deligne2007singularites,Maxim_lectures,pham}.\footnote{In particular, they find application in mirror symmetry for Fano varieties as they encode the Gromov--Witten invariants.\textcolor{orange}{add references---Maxim's mirror symmetry talk, Varchenko}.} In fact, the thimbles represent classes in the homology $H^{B,\theta}_N(X,f)$ which are relative to the preimage of $e^{i\theta}\infty$ under the funtion $f$, for a generic direction $\theta$.\footnote{The relative homology $H_i^{\theta}(X,f)$ is defined as the limit as $c\to e^{i\theta}\infty$, of $H_i^{\theta}(X,f^{-1}(S_c^+))$, where $S_c^+=\lbrace \zeta\in\C \vert \Re \zeta> c\rbrace$. Generic directions correspond to $\theta\neq \arg (\alpha-\beta)$. } Furthermore, if $f$ has non-degenerate critical points, they form a basis for $H_N^{B,\theta}(X,f)$. Then varying $\theta$, $H_N^{B,\theta}(X,f)$ forms a local system over $S^1=\R/2\pi \Z$ singular at $\theta=\arg(\alpha-\beta)$, and whose monodromy can be computed by Picard--Lefschetz formula (we refer to \cite[Section~1]{Arnold} and \cite[Section~3.3, Part II]{pham}). Equivalently, the monodromy data can be computed as the Stokes phenomena for Laplace transforms; indeed as a consequence of our Borel regularity result Theorem~\ref{thm:maxim}, thimble integrals can be turned into Laplace transforms.

Classical examples of thimble integrals (with polynomial $f$) are special functions, thus the Borel summability properties of their asymptotics can be studied either analytically (through the \textit{thimble projection reasoning} see the discussion in Section~\ref{sec:examples}) or geometrically (in terms of a certain Riemann--Hilbert problem \cite{Maxim_slide_ERC}\cite[Section 6.2]{kontsevich2022analyticity}).

In the $N$-dimensional analog of our set-up \ref{borel-reg:explanatory-power} the thimbles are actually steepest descent paths from the critical points; thus the asymptotic behaviour as $z\to e^{i\theta}\infty$ can be studied using the steepest descent method \cite{andersen2020resurgence,delabaere-howls,delabaere_dillinger_pham,Delabaere-Pham99,dingle1973asymptotic,Malgrange22,Pham83}.\footnote{In fact, analougus results hold when $f$ satisfies milder assumptions (see for instance~\cite[Section 1.2.2]{mistegard_phdthesis}).} However, the asymptotics of thimble integrals is typically a formal power series, and choosing a suitable resummation technique one should be able to recover the original function. This raises the question we address in this paper, namely to understand why Borel summation is effective for $1$-dimensional thimbles integrals. In a nutshell, our result follows from the fact that thimble integrals are generalized Laplace transforms, hence they are Borel regular functions. In addition, our approach wants to focus on the role of the position domain and it emphasizes how the geometry of the Laplace transform is the natural framework to describe Borel regularity for thimbles integrals (see Fenyes's lecture~\cite{Fenyes-ihes-lecture} and Section~\ref{sec:Laplace-Borel-general}).

@@

%\textcolor{orange}{hyperasymptotics \cite{Berry_Howls,Berry1991,Howls97,Howls}}
%\color{DarkTurquoise}
%These questions have been classically addressed by Pham \cite[Section 3.3]{pham}, Malgrange \cite[Theorem 6]{Malgrange22}, Berry and Howls \cite{Berry_Howls} and Sternin and Shatalov \cite[Theorem 3.9 and 3.10]{sternin1995borel}. In addition, the study of thimble integrals plays an important technical role in quantum mechanics, where infinite-dimensional exponential integrals are supposed to give the expectation values of observable quantities. Physicists often use Borel summation, resurgence and related techniques to assign values to these integrals \cite{costin_kruskal,pham1988resurgence}.
%\color{black}
%%Our approach wants to focus on the role of the position domain and it emphasizes how the geometry of the Laplace transform is the natural framework to describe Borel regularity for thimbles integrals (see Fenyes's lecture \cite{Fenyes-ihes-lecture}). In a nutshell, we use that thimble integrals are generalized Laplace transforms, to show they are indeed Borel regular functions.  
%     
\section{The Laplace and Borel transforms}\label{sec:Laplace-Borel-general}
\subsection{The geometry of the Laplace transform}
Classically, the Laplace transform turns functions on the position domain into functions on the frequency domain. In the study of Borel summation and resurgence, it's useful to see the position domain as a {\em translation surface} $B$, and the frequency domain as one of its cotangent spaces. Roughly speaking, the Laplace transform lifts holomorphic functions on $B$ to holomorphic functions on $T^* B$.
%
\subsubsection{Translation surfaces, briefly}
%
A translation surface is a Riemann surface $B$ carrying a holomorphic $1$-form $\lambda$~\cite{zorich2006flat}. A translation chart is a local coordinate $\zeta$ with $d\zeta = \lambda$. The standard metric on $\C$ pulls back along translation charts to a flat metric on $B$, with a conical singularity of angle $2\pi n$ wherever $\lambda$ has a zero of order $n-1 > 0$. We'll require $B$ to be finite-type and $\lambda$ to have a pole at each puncture. This kind of translation surface has a ``cylindrical end'' (see Figure \ref{fig:translation_surface}) at each puncture where $\lambda$ has order $-1$, and a ``$|2n|$-planar end'' (see Figure \ref{fig:translation_surface}) at each puncture where $\lambda$ has order $n-1 < -1$~\cite[Section 2.5]{gupta2013meromorphic} (or cite Aaron's article, which will hopefully present the same background in the translation surface context).
\begin{figure}[ht]
    \centering
    %\includegraphics{}
    \caption{On the left an example of translation surface with ``cylindrical end''. On the right, an example of translation surface with  ``$|2n|$-planar end''. }
    \label{fig:translation_surface}
\end{figure}
%
\subsubsection{Direction}\label{transl:dir}
%
The translation structure gives $B$ a notion of direction as well as distance. Away from the zeros of $\lambda$, which we'll call {\em branch points}, we can talk about moving upward, rightward, or at any angle, just as we would on $\C$. At a branch point of cone angle $2\pi n$, we can also talk about moving upward, rightward, or at any angle in $\R/2\pi\Z$, but here there are $n$ directions that fit each description. To make this more concrete, note that around any point $b \in B$, there's a unique holomorphic function $\zeta_b$ that vanishes at $b$ and has $d\zeta_b = \lambda$. \textcolor{VioletRed}{[If we define ``translation parameter'' earlier, we can say:] there's a unique translation parameter $\zeta_b$ that vanishes at $b$.} This function is a translation chart when $b$ is an ordinary point, and an $n$-fold branched covering when $b$ is a branch point of cone angle $2\pi n$. In either case, $\zeta_b \in e^{i\theta} [0, \infty)$ is a ray or a set of rays leaving $b$ at angle $\theta \in \R/2\pi\Z$.
%

@@

\item With respect to the flat metric, $f$ is uniformly of exponential type $\Lambda$ along $\Gamma_b^\theta$, and is locally integrable throughout.\footnote{Recall that a function is of exponential type $\Lambda$ if for every $\varepsilon>0$ there is a constant $A_\varepsilon$ (which depends on $\varepsilon$) such that $|f(x)|\leq A_\varepsilon e^{\Lambda+\varepsilon} $. We instead require that there exists a uniform constant $A$ such that for every $\varepsilon>0$ the bound holds.} Following \cite{reg-sing-volterra}, we require $f\in\singexp{\sigma}{\Lambda}(\Omega_b)$ with $\sigma>-1$.\footnote{The condition $\sigma>-1$ guarantees the integrability at $\zeta=b$.} The set $\Omega_b$ is a tubular neighbourhood of $\Gamma_b^\theta$. 

@@

\color{RoyalBlue}{$\singexpalg{\sigma}(\Omega_\alpha^\Lambda)$}
\color{Maroon}
Suppose $\Omega_\alpha$ is an open sector with $\zeta = \alpha$ at its tip, and an opening angle of $\pi$ or less. In this case, for any $\Lambda \in \R$, let \textcolor{magenta}{$\widehat{\Omega}_\alpha^\Lambda$} $\dualsector{\alpha}{\Lambda}$ be the union of the half-planes $\Re(e^{i\theta} z) > \Lambda$ over all angles $\theta$ in the opening of $\Omega_\alpha$. Then, let $\dualsingexp{\sigma}{\Lambda}(\Omega_\alpha)$ be the space of holomorphic functions $F$ on $\widehat{\Omega}_\alpha^\Lambda$ with $|F| \lesssim \Delta^\sigma$, where $\Delta$ is the distance to the boundary of $\widehat{\Omega}_\alpha^\Lambda$. The norm $\|F\|_{\sigma, \Lambda} = \sup_{\widehat{\Omega}_\alpha^\Lambda} \Delta^{-\sigma} |F|$ turns $\dualsingexp{\sigma}{\Lambda}(\Omega_\alpha)$ into a Banach space. \textcolor{orange}{Cite \cite{sternin1995borel} better for the theorem.}
\begin{prop}[following~\cite{sternin1995borel}]\label{prop:laplace-cont}
Let $\Omega_\alpha$ be an open sector with $\zeta = \alpha$ at its tip, and an opening angle of $\pi$ or less. For any $\sigma > 0$ and $\Lambda \ge 0$, and any angle $\theta$ in the opening of $\Omega_\alpha$, the Laplace transform $\laplace_{\zeta_\alpha, 0}^\theta$ is a continuous map $\singexp{\sigma-1}{\Lambda}(\Omega_\alpha) \to \dualsingexp{-\sigma}{\Lambda}(\Omega_\alpha)$, with a norm of at most $\Gamma(\sigma)$.
\end{prop}
\begin{proof}
Given some $f \in \singexp{\sigma-1}{\Lambda}(\Omega_\alpha)$, we compute
\begin{align*}
|\laplace_{\zeta_\alpha, 0}^\theta f| & = \left| \int_{\Gamma_{\zeta, \alpha}^\theta} e^{-z\zeta_\alpha} f\,d\zeta_\alpha \right| \\
& \le \int_{\Gamma_{\zeta, \alpha}^\theta} e^{-\Re(z\zeta_\alpha)} |\zeta_\alpha|^{\sigma-1} e^{\Lambda |\zeta_\alpha|} \|f\|_{\sigma-1, \Lambda}\,|d\zeta_\alpha| \\
& \le \int_{\Gamma_{\zeta, \alpha}^\theta} e^{(\Lambda - c_{z, \theta}|z|)|\zeta_\alpha|} |\zeta_\alpha|^{\sigma-1} \|f\|_{\sigma-1, \Lambda}\,|d\zeta_\alpha|,
\end{align*}
where $c_{z, \theta}$ is the cosine of $\arg(z) + \theta$. When $|z|$ is large and $c_{z, \theta}$ is positive, the integrand shrinks exponentially as $|\zeta|$ grows. This shows that for each angle $\theta$ in the opening of $\Omega_\alpha$, the integral defining $\laplace_{\zeta_\alpha, 0}^\theta f$ converges in some region of $\widehat{\Omega}_\alpha^\Lambda$. It also shows that for different angles $\theta$, the functions $\laplace_{\zeta_\alpha, 0}^\theta f$ match where their domains overlap. We can therefore glue these functions together into one big Laplace transform of $f$, defined at large values of $|z|$ across the whole opening angle of $\widehat{\Omega}_\alpha^\Lambda$.
